Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2022-01-22 FDT2020/response/response.tex

\response{Fixed.}
\item (10,14) it seems to me that the definition of the coordinates of $x^{(\ell+1)}$ other than $\ell+1$
should appear in the algorithm too, not just here.
\response{Line 3 in Algorithm 1 takes care of the other coordinates.}

@@

\item (11,14-25) I’m getting tripped up at line 17. I understand the case for coordinates which have $x_j=0$. But where do you deal with those that do not? I suppose I’m just
missing something obvious, but so could any other reader. Please clarify your proof here.
\response{Here the goal is to establish that there is a $z\leq x^{(\ell+1)}$. For the coordinates of $x^{(\ell+1)}$ that are zero we show the proof. The other coordinates of $x^{(\ell+1)}$ are 1. So we just need to emphasize that all the $\tilde{z}_i \leq 1$. This is added now, an a reference to an observation we added (Observation 7)}.

\item (12,10) “maintainS”  \response{Fixed.}
\item (13,28) “Growing and Pruning THE FDT tree”
\response{Fixed.}
\item (14,12) “It is easy to check” <– you do this quite a bit. Please go through your paper and
pay attention to all the “easy to check”, “obvious”, “clearly” etc. If these things really are easy, then just state them and don’t waste space. However, if some observation is necessary, think about the clearest and most concise way to impart this observation to the reader, and then do that instead of telling the reader that it should be “easy”.

\response{Thank you for the reminder. We find several occurrences of trivially, obviously, easy to check and clearly.}

\item (14,12) “THE set $L'$” \response{Fixed.}
\item (14,25) “From THE leaves” 
\response{Fixed.}
\item (16,11-17) You should discuss why the intuitive approach of binarizing an LP (just split each variable into two binary variables and copy the respective column of the
constraint matrix) does not lead to the same or better result. Or maybe it does?
\item (19,14) “1Mb cache” – I looked up this CPU and it has 1MB of L2 cache. Note that
1Mb $\neq$ 1MB, and “cache” is quite an ambiguous term (there are several types of CPU
cache). Moreover, the amount of cache is a property of the CPU and is usually not
stated, so I would just not comment on it. If you must, give a correct capacity and
indicate the level of the cache. \response{We removed the cache amount. Thanks.}
\item (19,22-26) Is there no good benchmark set for TAP? If not, please mention that as
the reason why you study random instances. (It is not good practise in experimental
algorithmics to study random instances unless necessary.)
\item (20,22) One idea to consider and comment on: did you warm-start each computation of FDT? It seems to me that you could reuse the model that FDT solves in each
iteration – basically, you only change the objective and fix the variables which corresponds to changing lower/upper bounds, so you should be able to save time on
building / destroying the model. I’m not sure how much the pruning model can be reused. Either way, please comment on this idea. The idea was successful elsewhere,
e.g. https://arxiv.org/abs/1911.08636 
\item (21,10) Christofides’ algorithm is no longer the fastest https://arxiv.org/abs/2007.01409

\response{We added a reference to the KKO paper in the footnote.}
\item (21,10-11) “In the classic graph-based…” <– this is a weird sentence; you should say
that you will re-cap the algorithm and then you can describe it. Right now it comes
out of nowhere.
\response{Rewrote this part: Let us remind our reader the Christofides algorithm for TSP: given a graph with edge costs find a minimum spanning tree of the graph. 
Next, we add an $O$-join, where $O$ is the set of odd-degree vertices in the spanning tree. Note that the solutions to the TSP are also 2-edge-connected, which means that Christofides algorithm also provides solutions to the 2EC.
The best-of-many Christofides (BOMC) algorithm~[AKS15] works better than the classic Christofides algorithm in practice[GW17].}

### 2022-01-22 FDT2020/Introduction.tex
finished minor comments - mostly

Formally, for some positive integer $k$, given a set of points $S\subseteq \mathbb{R}^k$ and a vector $x\in \mathbb{R}^k$, we say \textit{$x$ \underline{is} as a convex combination of points in $S$} if there exists a positive integer $\ell$ and $\theta \in [0,1]^\ell$, where $\sum_{i=1}^{\ell}\theta_i =1$ and $\tilde{x}^i\in S$ for $i\in [k]$ such that $\sum_{i=1}^{k}\theta_i \tilde{x}^i\leq x$. If a such a convex combination in polynomial time in the encoding size of the definition of $S$ and $x$, we say  \textit{$x$ \underline{can be written} as a convex combination of points in $S$}

@@

%\arash{I added this figure \ref{fig:CVpoint} but maybe we can put it later in the experiments section to make the introduction shorter?} \cindy{Yes that would be fine, if the reader could also use that figure then.  They could page forward to see the figure when reading the introduction.}
%\arash{I changed the paragraph below a little bit, can you check?} \cindy{Looks good. I just changed one sentence to the active voice.  If you're OK with that, you can delete these 2 comments. Thanks!}
Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ is achieved for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. Multiple groups of researchers have conjectured that $g(\2ec)\leq \frac{6}{5}$ (see \cite{alexander2006integrality,boydlegault,hn19}, but the only known upper bound on $g(\2ec)$ stands at $\frac{3}{2}$ \cite{wolsey}. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the cycle formed by the fractional edges, thereby confirming the conjecture for these instances. Note that the number of vertices in these instances can be arbitrarily high since the paths of edges with $x$-value 1 can be arbitrarily long.
\cindy{This sounds like a contribution of the paper.} \arash{Yes, particlarly since a similar research question have been studied by \cite{alexander2006integrality} and \cite{TSPcompute}}
\cindy{We probably need an explicit list of contributions.  It could be another subsection of the introduction.  This will help the reviewer and eventually, if published, the reader.}
\subsection{Contribution summary}
In summary, this papers's contributions are:
\begin{itemize}
\item We give an algorithm to express any feasible point for the LP relaxation of a binary IP
%(or a \{0,1,2\} IP) 
as a convex combination of feasible solutions, provided the IP integrality gap is bounded.  This is the Fractional Decomposition Tree (FDT) algorithm.
\item We give a simple  algorithm to give a feasible solution for any binary IP with bounded integrality gap. If this algorithm fails, then the integrality gap is infinite.
\item We experimentally show that FDT can give good approximate solutions to small instances of vertex-cover and tree augmentation problems.
\item We show that the integrality gap for 2EC (minimum 2-edge-connected sub-multigraph) is at most $\frac{6}{5}$ for Carr-Vempala points with at most $12$ vertices on the cycle formed by the fractional edges, matching the lower bound for integrality gap for 2EC for these instances.
\item We demonstrate (with 2EC) how FDT exeriments can support theoretical work on integrality gaps for specific problems. In particular, we computationally prove an upper bound on small instances of 2EC with a specific structure.
\end{itemize}

>>>>

Formally, for some positive integer $k$, given a set of points $S\subseteq \mathbb{R}^k$ and a vector $x\in \mathbb{R}^k$, we say \textit{$x$ \underline{is} as a convex combination of points in $S$} if there exists a positive integer $\ell$ and $\theta \in [0,1]^\ell$, where $\sum_{i=1}^{\ell}\theta_i =1$ and $\tilde{x}^i\in S$ for $i\in [k]$ such that $\sum_{i=1}^{k}\theta_i \tilde{x}^i\leq x$. If such a convex combination can be found in time polynomial in the encoding size of the definition of $S$ and $x$, then we say  \textit{$x$ \underline{can be written} as a convex combination of points in $S$}

@@

%\arash{I added this figure \ref{fig:CVpoint} but maybe we can put it later in the experiments section to make the introduction shorter?} \cindy{Yes that would be fine, if the reader could also use that figure then.  They could page forward to see the figure when reading the introduction.}
%\arash{I changed the paragraph below a little bit, can you check?} \cindy{Looks good. I just changed one sentence to the active voice.  If you're OK with that, you can delete these 2 comments. Thanks!}
Carr and Vempala \cite{Carr2004} showed that $g(\2ec)$ is achieved for instances where the optimal solution to $\min_{x\in \subtour(G)}cx$ is a Carr-Vempala point. Multiple groups of researchers have conjectured that $g(\2ec)\leq \frac{6}{5}$ (see \cite{alexander2006integrality,boydlegault,hn19}, but the only known upper bound on $g(\2ec)$ stands at $\frac{3}{2}$ \cite{wolsey}. We show that the integrality gap is at most $\frac{6}{5}$ for Carr-Vempala points with at most 12 vertices on the cycle formed by the fractional edges, thereby confirming the conjecture for these instances. Note that the number of vertices in these instances can be arbitrarily high since the paths of edges with $x$-value 1 can be arbitrarily long.
\cindy{This sounds like a contribution of the paper.} \arash{Yes, particlarly since a similar research question have been studied by \cite{alexander2006integrality} and \cite{TSPcompute}}
\cindy{We probably need an explicit list of contributions.  It could be another subsection of the introduction.  This will help the reviewer and eventually, if published, the reader.}
\subsection{Contribution summary}
In summary, this papers's contributions are:
\begin{itemize}
\item We give an algorithm to express any feasible point for the LP relaxation of a binary IP
%(or a \{0,1,2\} IP) 
as a convex combination of feasible solutions, provided the IP integrality gap is bounded.  This is the Fractional Decomposition Tree (FDT) algorithm.
\item We give a simple  algorithm to give a feasible solution for any binary IP with bounded integrality gap. If this algorithm fails, then the integrality gap is infinite.
\item We experimentally show that FDT can give good approximate solutions to small instances of vertex-cover and tree augmentation problems.
\item We show that the integrality gap for 2EC (minimum 2-edge-connected sub-multigraph) is at most $\frac{6}{5}$ for Carr-Vempala points with at most $12$ vertices on the cycle formed by the fractional edges, matching the lower bound for integrality gap for 2EC for these instances.
\item We demonstrate (with 2EC) how FDT exeriments can support theoretical work on integrality gaps for specific problems. In particular, we computationally prove an upper bound on small instances of 2EC with a specific structure.
\item We demonstrate (with VC) how FDT can be used to find feasible solutions of improved quality compared to those from the feasiblity pump \cite{fp1}.
\end{itemize}
